\documentclass{article}

\usepackage{mathtools,amsfonts}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{fullpage}

\begin{document}

\begin{center}
  \textbf{\Large Intermediate Test 1 Solutions}
  % LEVEL is Advanced, Intermediate or Beginner
  % NUMBER is the test number: 1, 2, etc.
  \\ \vspace{1em}
  \textbf{\large Stellenbosch Camp 2019}
\end{center}


\begin{enumerate}[1.]

\item[1.] % Emile, C
\textit{Find the number of paths from $A$ to $B$, where the only allowed moves are moving down along a diagonal line or moving left or right along a horizontal line, but never crossing the same line twice.}

\textit{Solution}:
Let us define the diagram in terms of layers, with $A$ being layer $1$, the two nodes below $A$ are layer $2$, etc. with $B$ as layer $7$.

Note that if you step down a layer, you can go to any of the other nodes on that layer in exactly $1$ way using lefts \& rights.
This implies that the number of paths to nodes in the same layer are the same, and are equal to the number of connecting branches between the previous and current layer mutliplied by the number of paths to a node on the previous layer.

$\implies$ The number of paths from $A$ to $B$ is:
$$1 \times 2 \times 4 \times 4 \times 4 \times 4 \times 2 = 2^{10}$$.
\vspace{6.81mm}

\item[2.] % PAMO 2016 Q1, G
\textit{Two intersecting circles, $C_1$ and $C_2$, have a common tangent which touches $C_1$ at $P$ and $C_2$ at $Q$. The two circles intersect at $M$ and $N$, where $N$ is nearer to $PQ$ than $M$ is. The line $PN$ meets the circle $C_2$ again at $R$. Prove that $MQ$ bisects $\angle PMR$.}

\textit{Solution}: Draw lines $PN$, $QN$ and $MN$. By the tan-chord theorem, we have that $\angle QPN=\angle PMN$ and $\angle PQN=\angle QMN$ and so $\angle PMQ=\angle PMN+\angle QMN=\angle QPN+\angle PQN$. Also $\angle RMQ=\angle RNQ$ since $RQNM$ is cyclic. From triangle $PNQ$, we obtain that $\angle QPN+\angle PQN= \angle RNQ$ hence $\angle PMQ=\angle RMQ$ and so $MQ$ bisects $\angle PMR$ 
\vspace{6.81mm}

\item[3.] % Emile, C
\textit{Emma and Emile play a game on a $2019 \times 2019$ board made up of unit grid squares. Emma plays first by placing a knight on one of the squares and thereafter they take turns to place a knight on a square that does not already contain a knight and is not attacked by one of the already placed knights. The first player who cannot do this loses. Can one of the players always guarantee that they will win? If so, which one? }

\textit{Solution}:
Emma can guarantee that she always wins.

Let us assume that the board is coloured like a chessboard. Emma moves first, so she can start by placing a knight on the block in the center of the board on her first move. 
Now, whatever move Emile plays, Emma plays the same move, but rotates it $180^o$ around the center of the board. 
If Emile makes a legal move, it will not invalidate the square rotated $180^o$ from it, since knights attack squares of opposite colour and the squares that are $180^o$ rotations of each other are the same colour.
This guarantees that Emma will always be able to play if Emile can play.
Therefore, the first person who will not be able to make a move is Emile, so Emma must win.
\vspace{6.81mm}

\item[4.] % Moldova 2018 8.6, A
\textit{Let $a$, $b$, and $c$ be positive real numbers such that $b, c \in [1,2)$ and
\[ \frac{a+b}{b(1+c)} +\frac{a+c}{c(1+b)} = 2. \]
Show that $a$, $b$, and $c$ are the lengths of the sides of a triangle.}

\textit{Solution}:
By multiplying out and simplifying the given equation, we get that $a = bc$.
Thus $a + b > a = bc \geq c$, and similarly $a+c > b$.
Finally,
\[ b+c > a = bc \iff 1 > 1 -b -c +bc = (1-b)(1-c), \]
which is true since $0 \leq b-1, c-1 < 1$.
\vspace{6.81mm}

\item[5.] % The Phil, N
\textit{Find the least positive integer $k$ such that $2050^{2051}$ can be written as a sum of $k$ $5$th powers.}

\textit{Solution}: 
By Fermat's Little Theorem: 
$$a^{10} \equiv _{11} 1 \; \forall a \in \mathbb{Z}, gcd(a, 11) = 1$$ 
$$\implies (a^5)^2 \equiv _{11} 1$$
$$\implies 11 \mid (a^5)^2 - 1$$
$$\implies 11 \mid (a^5 - 1)(a^5 + 1)$$
Since $11$ is prime, we must have $11 \mid a^5 - 1 \implies a^5 \equiv _{11} 1$ or $11 \mid a^5 + 1 \implies a^5 \equiv _{11} -1$. When $11 \mid a$, $a^5 \equiv _{11} 0$.
This shows that the only values of powers of $5$ mod $11$ are $-1$, $0$ and $1$. Notice that $2050^{2051} = (2050^{205})^{10} \times 2050 \equiv _{11} 4$ and the smallest way to express $4$ as the sum of some values from $\{-1, 0, 1\}$ is $4 = 1 + 1 + 1 + 1$.
This shows that $2050^{2051}$ cannot be expressed as the sum of less than $4$ fifth powers, so $k = 4$ is a lower bound.
$$(2050^{410} \times 2)^5 + (2050^{410} \times 2)^5 + (2050^{410} \times 1)^5 + (2050^{410} \times 1)^5 = (2050^{410 \times 5}) \times (2^5 + 2^5 + 1^5 + 1^5) = 2050^{2051}$$
This shows that $k = 4$ is possible, so $k = 4$ is the minimum.

\end{enumerate}

\end{document}
