\documentclass{article}

\usepackage{mathtools,amssymb}
\usepackage{enumerate}
\usepackage{fancyvrb}
\usepackage{fullpage}


\title{Advanced Test 1}
\author{Stellenbosch Camp 2019}
\date{Time: $2\frac{1}{2}$ hours}

\begin{document}
\maketitle
\thispagestyle{empty}


\begin{enumerate}[1.]

\item % Spanish MO Q1, C
A set $T$ of integers is called \emph{broken} if there are integers $a < b < c$ such that $a$ and $c$ are in $T$ but $b$ is not in $T$.

Find the number of broken subsets of $\{1, 2, \dotsc, 2019\}$.


\item % Venezuela K11 Q4, G
Let $ABC$ denote an equilateral triangle.
Let $M$ and $N$ denote the midpoints of $AB$ and $BC$, respectively.
Let $P$ be a point outside $ABC$ such that $APC$ is isosceles and right-angled at $P$.
Lines $PM$ and $AN$ meet at $I$.
Prove that $CI$ is the angle bisector of $\angle ACM$.


\item % British MO 2013-2014, N
A number written in base $10$ is a string of $3^{2019}$ digit $3$s.
No other digit appears.
Find the highest power of $3$ which divides this number.


\item % Irish 2018 Q7, C
There are $6$ eagles, $17$ snakes and $55$ mice in Wonderland.
An eagle can eat a snake or a mouse, but not another eagle.
A snake can eat a mouse, but not an eagle or another snake.
A mouse cannot eat an eagle, a snake or another mouse.

Whenever an eagle eats a snake, it turns into a mouse, and when it eats a mouse it turns into a snake. When a snake eats a mouse, it turns into an eagle.

After some time a situation is reached in which no animal can eat another animal.
What is the maximal possible number of animals alive in this situation?




\end{enumerate}


\vfill
% ASCII art
\centering
\begin{BVerbatim}
   .-"""""""-.    .'""."".""`.
  /   \ * /   \  /  ^ \   /^  \
 { ((@)\*/((@) }.{ ~Q '\ / Q ~}}
( `.~~ .V. ~~.'}.`._.'  V`._.'  )
( ( `-')  `-'  }   ~~     ~~    )
(  (   ) ^ ^   0 "  " "   (  (   )
(      )  ^^  00   "      (      .)
( (   )     ^ 00     "    ( (   ) )
 {   )  ^^   .0'0  "   "   (     }
  { }  ^   ))    0          (   }
   `--www--'       "--mmm------ "    
\end{BVerbatim}

\end{document}
