\documentclass{article}

\usepackage{amsmath,amssymb}
\usepackage{fullpage}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{graphicx}


\begin{document}

\begin{center}
	\textbf{\Large Beginner Monthly Problem Set}
	\\ \vspace{1em}
	\textbf{\large Due: Friday, 17 January 2020}
\end{center}

\begin{enumerate}[1.]

\vspace{6pt}
\item % 2010 Team Q6 
How many positive square numbers are factors of 1600?

\vspace{6pt}
\item % 2011 Team Q5
A palindrome is a positive integer whose digits are the same when read forwards or backwards.
For example, 2882 is a four-digit palindrome and 49194 is a five-digit palindrome. There are
pairs of four-digit palindromes whose sum is a five-digit palindrome. One such pair is 2882
and 9339. How many such pairs are there including this pair?

\vspace{6pt}
\item %2013 Team Q6
Show all the ways the number 365 can be written as the sum of two or more different perfect
square numbers?

\vspace{6pt}
\item %2014 Team Q5
How many squares on an $8 \times 8$ chessboard are more black than white?

\vspace{6pt}
\item %2014 Team Q8
In triangle $ABC$, $\widehat{A} = 45^{\circ}$. Point $P$ is on $BC$, $Q$ on $AB$ and $R$ on $AC$ such that $BP = QP$ and $CP = RP$. Find $Q \widehat{P} R$.

\vspace{6pt}
\item % 2016 Team Q3
If $X$ and $Y$ are two digits and the five digit number $30XY5$ can be expressed as the product $225 \times n$, find all possible positive integer values of $n$.

\vspace{6pt}
\item %Junior 2009 Team Competition
The symbol $\left \lfloor{x}\right \rfloor$ means the greatest integer less than or equal to $x$. For example, 
\begin{center}
	$\left \lfloor{5.7}\right \rfloor = 5$ and $\left \lfloor{4}\right \rfloor = 4$.
\end{center} 
Calculate the value of the sum
\begin{center}
	$\left \lfloor{\sqrt{1}}\right \rfloor$ + $\left \lfloor{\sqrt{2}}\right \rfloor$ + $\left \lfloor{\sqrt{3}}\right \rfloor$ + $\left \lfloor{\sqrt{4}}\right \rfloor$ + ... + $\left \lfloor{\sqrt{49}}\right \rfloor$ + $\left \lfloor{\sqrt{50}}\right \rfloor$
\end{center}

\vspace{6pt}
\item %2012 Team Q7
On a page, there are 10 circles. Any pair of them intersect at exactly 2 points but no 3 of
them intersect at one common point. Into how many parts do these circles divide the page?

\vspace{6pt}
\item %2015 Team Q9
Twelve points are equally spaced on the circumference of a circle. How many different rectangles may be formed by joining any four points at a time?

\vspace{6pt}
\item %2012 Team Q5
What is the remainder when $6^{2012} + 8^{2021}$ is divided by 49?

\end{enumerate}

\vspace{8pt}
\textbf{\Large Email submission guidelines}
\begin{itemize}
	\item Email your solutions to \href{mailto:samf.training.assignments@gmail.com}{\texttt{samf.training.assignments@gmail.com}}.
	\item In the subject of your email, include your name and the level of the assignment (Beginner, Intermediate or Senior).
	\item Submit each question in a single separate PDF file (with multiple pages if necessary), with your name and the question number written on each page.
	\item If you take photographs of your work, use a document scanner such as Office Lens to convert to PDF.
	\item If you have multiple PDF files for a question, combine them using software such as PDFsam.
\end{itemize}

\end{document}
